---
title: "Reproducible Research Documentation"
author: "Craig Tedeton"
date: "May 22, 2016"
output: html_document
---
### Code for reading in the dataset and/or processing the data
```{r echo = TRUE}
library(readr) ; library(ggplot2) ; library(plyr) ; library(lattice)
# Project Code: Section 1
# Getting/organizing the data
fileUrl <- "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"
download.file(fileUrl, "./data/repdata%2Fdata%2Factivity.zip")
unzip("./data/repdata%2Fdata%2Factivity.zip", exdir = "./data")
fpath <- file.path("./data")
filelist <- list.files(fpath, recursive = TRUE)
filelist
fileName <- "./data/activity.csv"

# Reading/reviewing of data source
activ <- read.csv(fileName)
summary(activ)
rows <- nrow(activ) ; cols <- ncol(activ)
rows ; cols

# convert date field to date type with only weekdays
activ$day <- weekdays(as.Date(activ$date))
activ$DateTime<- as.POSIXct(activ$date, format="%Y-%m-%d")

myCleanData <- activ[!is.na(activ$steps),]

# mean number of steps take per day/date
sumData <- aggregate(activ$steps ~ activ$date, FUN = sum, )
colnames(sumData)<- c("Date", "Steps")
```

### Histogram of the total number of steps taken each day
```{r echo = TRUE}
# plot data to histogram of steps per day/date
hist(sumData$Steps, breaks = 5, xlab = "Steps", main = "Total Steps per Day")
```

### Mean and median number of steps taken each day
```{r echo = TRUE}
## Mean seps by day
as.integer(mean(sumData$Steps))

# Median steps by day
as.integer(median(sumData$Steps))

# avg number of steps per interval
interData <- ddply(myCleanData, .(interval), summarize, Avg = mean(steps))
```

### Time series plot of the average number of steps taken
```{r echo = TRUE}
# plot of avg number of steps per interval
p <- ggplot(interData, aes(x = interval, y = Avg), xlab = "Interval", ylab = "Avg Number of Steps")
p + geom_line() + xlab("Interval") + ylab("Avg Number of Steps")+ggtitle("Avg Number of Steps per Interval")
```

### The 5-minute interval that, on average, contains the maximum number of steps
```{r echo = TRUE}
# maximum steps by interval
maxStep <- max(interData$Avg)

# interval containing  maximum avg number of steps
interData[interData$Avg == maxStep, 1]
```

### Code to describe and show a strategy for imputing missing data
```{r echo = TRUE}
##Number of NAs in original data set
nrow(activ[is.na(activ$steps),])

## Create the average number of steps per weekday and interval
avgData <- ddply(myCleanData, .(interval, day), summarize, Avg = mean(steps))

## Create dataset with all NAs for substitution
naData<- activ[is.na(activ$steps),]
## Merge NA data with average weekday interval for substitution
newData <- merge(naData, avgData, by = c("interval", "day"))

## Reorder the new substituded data in the same format as clean data set
newData2<- newData[,c(6,4,1,2,5)]
colnames(newData2)<- c("steps", "date", "interval", "day", "DateTime")

# merge the NA averages and non NA data
mergedData <- rbind(myCleanData, newData2)

# sum of steps per day to compare with step 1
sumData2 <- aggregate(mergedData$steps ~ mergedData$date, FUN = sum, )
colnames(sumData2)<- c("Date", "Steps")

# mean of Steps with NA data handled
as.integer(mean(sumData2$Steps))

# median of Steps with NA data handled
as.integer(median(sumData2$Steps))
```

### Histogram of the total number of steps taken each day after missing values are imputed
```{r echo = TRUE}
# histogram of total steps per day, by data set to show impact
hist(sumData2$Steps, breaks = 5, xlab = "Steps", 
     main = "Total Steps per Day with NAs Fixed", col = "Blue")
hist(sumData$Steps, breaks = 5, xlab = "Steps", 
     main = "Total Steps per Day with NAs Fixed", col = "Dark Grey", add = T)
legend("topleft", c("Imputed Data", "Non-NA Data"), fill = c("Blue", "Dark Grey") )
```

### Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends
```{r echo = TRUE}
# new category based on days of the week
mergedData$DayCategory <- ifelse(mergedData$day %in% c("Saturday", "Sunday"), "Weekend", "Weekday")

# summary data by interval and type of day
interData2 <- ddply(mergedData, .(interval, DayCategory), summarize, Avg = mean(steps))

# lot data in a panel plot
xyplot(Avg ~ interval|DayCategory, data = interData2, type = "l",  layout = c(1,2),
       main="Avg Steps per Interval Based on Type of Day", 
       ylab="Avg Number of Steps", xlab = "Interval")
```

### All of the R code needed to reproduce the results (numbers, plots, etc.) in the report
```{r ProjectCode, echo = TRUE}

```


